Near Optimal Behavior with Approximate State Abstraction

### 2016-02-03 paper/theory_results.tex
Lemma 3 and 4 proofs added

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{mult}}$ and $\ep_{\text{bolt}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V_G^{\pi^*_G}(s) - V_G^{\pi^*_{A}}(s) | \leq poly(\epsilon).
\end{equation}
\end{thm}

@@

% Subsection: Boltzmann over Optimal Q
\enote{Motivate Boltzmann + multinomial}

@@

Similarly to the multinomial abstraction, we let $\phi_{bolt}$ define a type of abstraction that, for fixed $\epsilon$ satisfies:
\begin{multline}
\ep_{bolt}(s_1) = \ep_{bolt}(s_2) \rightarrow \\
\forall_{a} \left|\frac{e^{Q^*(s_1,a)}}{\sum_b e^{Q^*(s_1,b)}} - \frac{e^{Q^*(s_1,a)}}{\sum_b e^{Q^*(s_1,b)}}\right| \leq \epsilon
\end{multline}

{\bf Lemma 4:} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some constant positive $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq 
\epsilon:
\end{equation}

@@

\dnote{Proof sketch}

\dnote{(Probably put full proofs in appendix if we have room?}

@@

We let $\phi_{mult}$ define a type of abstraction that, for fixed $\epsilon$ satisfies:
\begin{multline}
\ep_{mult}(s_1) = \ep_{mult}(s_2) \rightarrow \\
\forall_{a} \left|\frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)} - \frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)}\right| \leq \epsilon
\end{multline}

\dnote{We also need to add the point about the denominator bound...}

{\bf Lemma 3:} When $S_A$ is created using a function of the $\ep_{mult}$ type, for some constant positive $k \in \mathbb{R}$:
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{\left(\textsc{QMax}|\mathcal{A}| + k \epsilon + k \epsilon^2\right)}{1-\gamma}:
\end{equation}

@@

% Possibly a lemma for NP-Completeness of the minimal such phi for all abstractions (was a result for model in previous cases).

>>>>

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{bolt}}$ and $\ep_{\text{mult}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V_G^{\pi^*_G}(s) - V_G^{\pi^*_{A}}(s) | \leq poly(\epsilon).
\end{equation}
\end{thm}

@@

@@

Here we introduce the approximate abstraction family, $\phi_{bolt}$, which aggregates states with similar Boltzmann distributions on optimal Q values as equivalent. This family of abstractions is appealing as the Boltzmman distribution over Q-values has been of interest in the reinforcement learning literature due to its natural balancing of exploration and exploitation~\cite{sutton1998reinforcement}.

We let $\phi_{bolt}$ define a family of abstractions that, for fixed $\epsilon$ satisfies:
\begin{multline}
\ep_{bolt}(s_1) = \ep_{bolt}(s_2) \rightarrow \\
\forall_{a} \left|\frac{e^{Q^*(s_1,a)}}{\sum_b e^{Q^*(s_1,b)}} - \frac{e^{Q^*(s_2,a)}}{\sum_b e^{Q^*(s_2,b)}}\right| \leq \epsilon
\label{eq:phi_bolt}
\end{multline}

We also assume that the difference in normalizing terms is bounded by some non-negative constant, $k\in \mathbb{R}$, of $\epsilon$:
\begin{equation}
\sum_b e^{Q^*(s_1,b)} - \sum_b e^{Q^*(s_2,b)} \leq k*\epsilon
\label{eq:bolt_denom}
\end{equation}

\begin{lma} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some non-negative constant $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon\left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k\right)}{(1-\gamma)^2}
\end{equation}
\label{lma:bolt_lemma}
\end{lma}

We prove Lemma~\ref{lma:bolt_lemma} by using the approximation for $e^x$:
\begin{equation}
1 + x + \delta = e^x
\label{eq:e_to_x_approx}
\end{equation}
Where $\delta$ is the error associated with the approximation.

@@

By the approximation in Equation~\ref{eq:e_to_x_approx} and the assumption in Equation~\ref{eq:bolt_denom}:
\begin{align*}
|\frac{1 + Q(s_1,a) + \delta_1}{\sum_j e^{Q(s_1,a_j)}} - \frac{1 + Q(s_2,a) + \delta_2}{\sum_j e^{Q(s_1,a_j)} \pm k\epsilon}| \leq \epsilon
\end{align*}

It follows that:
\begin{equation}
|Q(s_1,a) - Q(s_2,a)| \leq \epsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k \right)
\label{eq:bolt_qs}
\end{equation}

Consequently, we can consider $\widetilde{\phi}_{bolt}$ as a type of $\epQ$, where the difference in Q values is bounded by: $\epsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k \right)$. The Lemma follows directly from Lemma~\ref{lma:Q*}.

@@

We consider a multinomial distribution over $Q^*$ for similar reasons as the Boltzmann. We let $\phi_{mult}$ define a type of abstraction that, for fixed $\epsilon$ satisfies.
\begin{multline}
\ep_{mult}(s_1) = \ep_{mult}(s_2) \rightarrow \\
\forall_{a} \left|\frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)} - \frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)}\right| \leq \epsilon
\end{multline}

We also assume that the difference in normalizing terms is bounded by some positive factor, $k \in \mathbb{R}$, of $\epsilon$:
\begin{equation}
|\sum_i Q(s_1,a_i) - \sum_j Q(s_2,a_j) | \leq k\epsilon
\end{equation}

\begin{lma} When $S_A$ is created using a function of the $\ep_{mult}$ type, for some non-negative constant $k \in \mathbb{R}$:
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{\frac{2\epsilon|\mathcal{A}|}{1-\gamma} + k \epsilon^2 + k}{(1-\gamma)^2}
\end{equation}
\end{lma}

@@

The proof follows a near identical strategy to that of Lemma~\ref{lma:bolt_lemma}.

\dnote{Bounds for multi and bolt suspisiously similar. Double check math}
